---
title: "Multisectoral Gravity Model"
---
### /!\ Read Carefully the following:
### For the baseline results (i.e., with Y60 as maritime flows), the replicator is expected to change the working directory in line 23: this is the only change required to run this do-file on any computer.
### For most alternative scenarios (Y50, Y70, Y0, as well as Y60 with 'low' and 'high' sigma), the replicator must also change the sigma defined in line 28 accordingly.
### For the alternative scenario in which all betas are included (not only the significant ones), the replicator must change the definition of beta in line 69-70 accordingly. Note that the replicator can also run the estimations using alternative levels of significance if wanted.
### In our study, we estimated the impact of a $40/tCO2 carbon tax. However, the replicator can simulate any carbon tax by simply replacing 40 with the desired value in line 35. Similarly, if one wants to simulate the impact of a carbon tax in a different reference year (i.e., different from 2019), one simply needs to replace the variable pt defined in line 36 with the annual HFO price value for the desired reference year.

### Preliminary Steps and Importation of Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) #For importation and exportation of data
library(foreign) #For exportation of data in .dta format
library(tidyr) #For computations
library(dplyr) #For computations
library(tensor) #For computations
```

### Importation of Data and Selection of Input Parameters: These parameters can be changed for the alternative scenarios
```{r}
setwd('')
data = read_dta("Results/Estimation Results.dta")
K = length(unique(data$Product))
N = length(unique(data$Exporter))

sigma <- data$sigma_60 #Baseline value of sigma to use for the main results
#sigma <- data$sigma_50 #Alternative value of sigma to use if the replicator defined 'local = 50' in the previous "03_PPML_Estimations.do"
#sigma <- data$sigma_70 #Alternative value of sigma to use if the replicator defined 'local = 70' in the previous "03_PPML_Estimations.do"
#sigma <- data$sigma_0 #Alternative value of sigma to use if the replicator defined 'local = 0' in the previous "03_PPML_Estimations.do"
#sigma <- data$sigma_60_low #Alternative value of sigma to use if the replicator wants to test for the first quartile sigma
#sigma <- data$sigma_60_high #Alternative value of sigma to use if the replicator wants to test for the third quartile sigma

tax <- 40 * 3.12 #We use 3.12 as the coefficient between USD/tCO2 and USD/tFuel. To simulate the tax, one can replace 40 with the desired value of the carbon tax (in USD per ton of CO2)
pt <- 377.4167   #We include the value of HFO in 2019 (can be changed if using another value of Fuel prices or another year of reference)
```

### Verification that all Yijk are (not strictly) positive
```{r}
positives <- all(data$Yijk >= 0)
if (positives) {
  print("All Yijk are positive or null.")
} else {
  print("Error: There is at least one negative Yijk.")
}
```

### Initialization of Arrays and Matrixes
```{r}
data$Exporter <- as.factor(data$Exporter)
data$Importer <- as.factor(data$Importer)
data$Product <- as.factor(data$Product)

N <- length(levels(data$Exporter)) #Nb of exporters = Nb of importers = N
K <- length(levels(data$Product))  #Nb of sectors = K

Yijk <- array(NA, dim = c(N, N, K))
beta <- array(NA, dim = c(N, N, K))
Seadistance <- array(NA, dim = c(N, N, K))
tijk_hat <- array(NA, dim = c(N, N, K))

for (i in 1:nrow(data)) {
  Exporter <- as.integer(data$Exporter[i])
  Importer <- as.integer(data$Importer[i])
  Product <- as.integer(data$Product[i])
  
  Yijk[Exporter, Importer, Product] <- data$Yijk[i]
  beta[Exporter, Importer, Product] <- data$beta[i] * data$significant05[i] #We select only betas significantly different from 0
  #beta[Exporter, Importer, Product] <- data$beta[i]                        #If one wants to keep all betas and not only the significant ones
  Seadistance[Exporter, Importer, Product] <- data$Seadistance[i]
}
for (i in 1:N) {
 for (j in 1:N) {
  for (k in 1:K) {
     tijk_hat[i,j,k] <- ( ( 1 + Seadistance[i,j,k] * (pt + tax)) / (1 + Seadistance[i,j,k] * pt) ) ^ (beta[i,j,k] / (1 - sigma[k]))
    }
  }
}
```

### Preliminary Computations
```{r}
preliminary_computations = function(Yijk, sigma) {
  N <- dim(Yijk)[1]
  K <- dim(Yijk)[3]

  Yik <- apply(Yijk, c(1, 3), sum) #Yik: Sum of Yijk over j (Matrix N*K)
  Yi <- apply(Yik, 1, sum) #Yi: Sum of Yijk over j and k (Vector 1*N)
  Y <- sum(Yi) #Y: Sum of Yijk over i, j, and k (Constant)

  Ejk <- apply(Yijk, c(2, 3), sum) #Ejk: Sum of Yijk over i (Matrix N*K)
  Ej <- apply(Ejk, 1, sum) #Ej: Sum of Yijk over i and k (Vector 1*K)

  lambda <- array(dim = c(N, N, K)) #λijk=Yijk/Ejk (Array N*N*K)
    for (i in 1:N) {
     for (j in 1:N) {
      for (k in 1:K) {
              lambda[i,j,k] <- Yijk[i,j,k] / Ejk[j,k]
      }
     }
    }

  phi <- Ej / Yi #φi=Ei/Yi (Vector 1*N)
  alpha <- Ejk / Ej #αik=Eik/Ei (Matrix N*K)

  return(list("Yik" = Yik, "Yi" = Yi, "Y" = Y, "Ejk" = Ejk, "Ej" = Ej, "lambda" = lambda, "phi" = phi, "alpha" = alpha))
}

  preliminary_results <- preliminary_computations(Yijk, sigma)
```

### Verification that all Yik and all Ejk are strictly positive
```{r}
positives <- all(preliminary_results$Yik > 0)
if (positives) {
  print("All Yik are strictly positive.")
} else {
  print("Error: There is at least one Yik equal to zero")
}

positives <- all(preliminary_results$Ejk > 0)
if (positives) {
  print("All Ejk are strictly positive.")
} else {
  print("Error: There is at least one Ejk equal to zero")
}
```

### Implementation of the Loop
```{r}
iterative_loop <- function(preliminary_results, sigma, tijk_hat, epsilon = 0.001, max_iterations = 1000) {
    Yik <- preliminary_results$Yik
    Ejk <- preliminary_results$Ejk
    lambda <- preliminary_results$lambda
    alpha <- preliminary_results$alpha
    phi <- preliminary_results$phi
    Y <- preliminary_results$Y

    N <- dim(Yik)[1]
    K <- dim(Yik)[2]

    Pjk_hat <- array(1, dim = c(N, K))
    Ejk_prime <- Ejk
    Yik_hat <- array(1, dim = c(N, K))
    Yik_tilde <- array(1, dim = c(N, K))
    
    intermediate_array <- array(dim = c(N, N, K))
    intermediate_matrix <- array(dim = c(N, K))
    intermediate_vector <- numeric(N) # dimension = N

    convergence <- FALSE

    for (iteration in 1:max_iterations) {
        Yik_hat_prev <- Yik_hat
        
        # We estimate Yik(n)_tilde
        for (i in 1:N) {
          for (j in 1:N) {
            for (k in 1:K) {
        intermediate_array[i,j,k] <- lambda[i,j,k] * (tijk_hat[i,j,k] ^ (1 - sigma[k])) * Ejk_prime[j,k] / (Pjk_hat[j,k] ^ (1 - sigma[k]))
            }
          }
        }
        intermediate_matrix <- apply(intermediate_array, c(1,3), sum)
        for (i in 1:N) {
          for (k in 1:K) {
            Yik_tilde[i,k] <- (intermediate_matrix[i,k] / Yik[i,k]) ^ (1 / sigma[k])
          }
        }      
        
        # We estimate Yik(n)_hat
        intermediate_matrix <- Yik_tilde * Yik
        intermediate_scalar <- sum(intermediate_matrix)
        for (i in 1:N) {
          for (k in 1:K) {
            Yik_hat[i, k] <- Yik_tilde[i, k] * Y / intermediate_scalar
          }
        } 
        
        # We estimate Pjk(n)_hat
        for (i in 1:N) {
          for (j in 1:N) {
            for (k in 1:K) {
        intermediate_array[i,j,k] <- lambda[i,j,k] * (Yik_hat[j,k] * tijk_hat[i,j,k]) ^ (1 - sigma[k])
            }
          }
        }
      intermediate_matrix <- apply(intermediate_array, c(2, 3), sum)
      for (j in 1:N) {
          for (k in 1:K) {
             Pjk_hat[j, k] <- intermediate_matrix[j,k] ^ (1 / (1-sigma[k]))
          }
        }    
        
        # We estimate Ejk(n)_prime
        intermediate_matrix <- Yik_hat * Yik
        intermediate_vector <- apply(intermediate_matrix, 1, sum)
        for (j in 1:N) {
            for (k in 1:K) {
                Ejk_prime[j, k] <- alpha[j, k] * phi[j] * intermediate_vector[j]
            }
        }

        # We check if convergence is achieved
        if (any(is.na(Yik_hat), na.rm = TRUE) || any(is.na(Yik_hat_prev), na.rm = TRUE)) {
            warning("NA values found in Yik_hat")
            break
        }
         if (max(abs((Yik_hat - Yik_hat_prev)/Yik_hat_prev), na.rm = TRUE) <= epsilon) {
            convergence <- TRUE
            break
         }
    }

    return(list("Pjk_hat" = Pjk_hat, "Ejk_prime" = Ejk_prime, "Yik_hat" = Yik_hat, "Yik_tilde" = Yik_tilde, "convergence" = convergence))
}

iterative_results <- iterative_loop(preliminary_results, sigma, tijk_hat)
```

### Final Computations
```{r}
final_computations <- function(iterative_results, preliminary_results, sigma) {
  N <- dim(preliminary_results$Yik)[1]
  K <- dim(preliminary_results$Yik)[2]
  Ejk_prime <- iterative_results$Ejk_prime
  Ejk <- preliminary_results$Ejk
  Yik_hat <- iterative_results$Yik_hat
  Pjk_hat <- iterative_results$Pjk_hat
  lambda <- preliminary_results$lambda
  Enik_prime <- iterative_results$Enik_prime
  alpha <- preliminary_results$alpha
  Ej <- preliminary_results$Ej

  Wi_hat <- numeric(N)  
  Pi_hat <- numeric(N)  
  Yijk_prime <- array(dim = c(N, N, K))  
  lambda_hat = array(dim = c(N, N, K))
  intermediate_array <- array(dim = c(N, N, K))
  intermediate_matrix <- array(dim = c(N, K))

  # We estimate Ei_prime
  Ei_prime = apply(Ejk_prime, 1, sum)
  
  # We estimate lambda_hat
  for (i in 1:N) {
   for (j in 1:N) {
    for (k in 1:K) {
     intermediate_array[i,j,k] <- lambda[i,j,k] * (Yik_hat[i,k] * tijk_hat[i,j,k])^(1-sigma[k])
    }
   }
  }
  intermediate_matrix <- apply(intermediate_array, c(2,3), sum)
  for (i in 1:N) {
   for (j in 1:N) {
    for (k in 1:K) {
     lambda_hat[i,j,k] <- ((Yik_hat[i,k] * tijk_hat[i,j,k])^(1-sigma[k])) / intermediate_matrix[j,k]
    }
   }
  }
  
  # We estimate Pi_hat
  for (i in 1:N) {
   for (k in 1:K) {   
    intermediate_matrix[i,k] <- Pjk_hat[i,k]^alpha[i,k]
   }
  }
  Pi_hat <- apply(intermediate_matrix, 1, prod)
  
  # We estimate Wi_hat
  for (i in 1:N) {
   Wi_hat[i]= Ei_prime[i]/(Ej[i]*Pi_hat[i])
  }
  
  # We estimate Yijk_prime
  for (i in 1:N) {
   for (j in 1:N) {
    for (k in 1:K) {
     Yijk_prime[i,j,k] <- Yijk[i,j,k] * lambda_hat[i,j,k] * Ejk_prime[j,k] / Ejk[j,k]
    }
   }
  }
  
  list(Wi_hat = Wi_hat, Pi_hat = Pi_hat, Yijk_prime = Yijk_prime)
}
final_results <- final_computations(iterative_results, preliminary_results, sigma)
```

### Storage and Export of Results
```{r}
Yijk_prime <- aperm(final_results$Yijk_prime, c(3,2,1)) # We reorganize the array to ensure the good position of each cell in the data.frame
Yijk_prime_vector <- as.vector(Yijk_prime)
Yijk_prime_df <- data.frame(Yijk_prime = Yijk_prime_vector)
data$Yijk_prime <- Yijk_prime_df$Yijk_prime

Wi_hat <- final_results$Wi_hat
Pi_hat <- final_results$Pi_hat
Wi_Pi_df <- data.frame(Wi_hat = rep(Wi_hat, each = length(unique(data$Importer)) * length(unique(data$Product))),
                        Pi_hat = rep(Pi_hat, each = length(unique(data$Importer)) * length(unique(data$Product))))  
data <- cbind(data, Wi_Pi_df)
head(data)

write_dta(data, "Results/Final Results.dta")
```
